### 2019-08-02 README.Rmd
fix link in readme

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
```

# Resources for Morey & Hoekstra (2019) <img src="man/figures/logo.png" align="right" alt="" width="120" />

Data, analysis code, and demonstrations for Morey & Hoekstra's (2019) statistical cognition experiment, which was run over the Christmas season in 2018. You can find more details at https://richarddmorey.github.io/Morey_Hoekstra_StatCognition/.

* [Installation](#installation)
* [Task demonstration](#task-demonstration)
* [Shiny app](#shiny-app)

## Installation

If you're using a recent version MacOS, you'll need to install [Xquartz](https://www.xquartz.org/index.html) for the fonts.

To fully use the package, you'll probably need to install several packages that are not on CRAN:

``` r
# install.packages('devtools') # if needed
devtools::install_github("crsh/papaja")
devtools::install_github("rstudio/gt")
```

Then you can install this package with

``` r
devtools::install_github("richarddmorey/Morey_Hoekstra_StatCognition", build_vignettes = TRUE, depedencies = TRUE)
```

## Task demonstration

A short video showing the task can be found under [`inst/manuscript/img/santa_task_example.mp4`](https://github.com/richarddmorey/Morey_Hoekstra_StatCognition/blob/master/inst/manuscript/img/santa_task_example.mp4)  (click 'View raw' to download).

You can find an [online version of the task demonstration](https://richarddmorey.github.io/Morey_Hoekstra_StatCognition/articles/task_demo.html).

Users of the package can also load a version of the experimental task from within R:

``` r
task_demo()
```

## Shiny app

Running the included shiny app will allow you to explore individual participants' response patterns.

``` r
explore_data()
```
You can access an already-deployed version of the shiny app at https://richarddmorey.shinyapps.io/explore/.

The app will let you:

* Search, sort, and filter the data set by experimental condition, response, etc.
* See the selected participant's interface at the poiint of decision (both experimental samples and random shuffle reports)
* See an animation showing the selected participant's behaviour over time, up to the point of decision
* Read the selected participant's responses to the questions about their strategy
* Search the strategy text using [POSIX regular expressions](https://stat.ethz.ch/R-manual/R-devel/library/base/html/regex.html), optionally (try "chance" to search for all instances of "chance"; "color|colour" will search for mentions of either of the two spelling variants of "color")